### 2021-08-23 Bernoulli/AnswerReviews.tex
Added answer to referees

\documentclass[12pt]{amsart}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage[round]{natbib}
\usepackage{amssymb,amsmath,latexsym,xcolor}

%\textwidth  14.5cm
%\textheight 20cm
%\voffset=-1cm
%\hoffset=-1cm

\newcommand{\rref}[1]{{\small \textcolor{gray}{#1}} }
\newcommand{\cl}[1]{ \textcolor{blue}{#1}}

\begin{document}

\begin{center}
\textbf{Answer to Referees for \textit{"A pseudo-marginal sequential Monte Carlo online smoothing algorithm"}}
\end{center}

\bigskip

We would like to thank the associate editor and the referees for the attention they paid to our work and their positive feedbacks.

\bigskip

Minor errors and typos listed by R2 were corrected as suggested. Pkease find below a few detailed answers.

\medskip

{\bf p.7 Just before 2.2.2, it would be nice to comment on where one might get suitable adjustment weights $\vartheta_n$ from. How are these to be chosen or tuned?}\\
{\em }

\medskip

{\bf p.29 Appendix A in the supplement has in the numerator that arises from . I think this should read instead?}\\
{\em The normalizing term given in the appendix is correct, see (2.10).}

\medskip

{\bf p.30 Appendix A has in the denominator for  which I think should read  as there is a dangling dxn+1 otherwise.}\\
{\em The denominator is correct the definition of $\mathsf{X}^n$ given in the appendix is $\mathsf{X}^n = \mathsf{X}_0 \times \ldots \times  \mathsf{X}_n$.}

\medskip

\end{document}